---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RTTshiny

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/nhs-bnssg-analytics/RTT_compartmental_modelling/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nhs-bnssg-analytics/RTT_compartmental_modelling/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/nhs-bnssg-analytics/RTT_compartmental_modelling/graph/badge.svg?token=O2KBKLVUOD)](https://codecov.io/gh/nhs-bnssg-analytics/RTT_compartmental_modelling)
<!-- badges: end -->

The goal of RTTshiny is to provide an interface to a stock-and-flow model on NHS Referral to Treatment data, to allow planners to understand how changes in future clock stops can effect the waiting list size and knock on performance.

## Run without installing

If you have R installed on your machine, the following command may work to run the application without installing the package:

``` r
# install.packages("shiny")
shiny::runGitHub("RTT_compartmental_modelling", "nhsengland")
```

## Installation

You can install the development version of RTTshiny from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github(
  "nhsengland/RTT_compartmental_modelling@dev",
  dependencies = "Suggests"
)
```

## Usage

To launch the Shiny app following installation:

```{r}
#| label: "usage"
#| eval: false
RTTshiny::run_app()
```


## Web pages

See more information about the model on the [pkgdown website](https://nhs-bnssg-analytics.github.io/RTT_compartmental_modelling/).
